\documentclass[10pt,twocolumn]{article}
\usepackage{amsmath, amssymb, booktabs, geometry, titlesec, parskip}
\usepackage{graphicx}
\usepackage{url}
\usepackage{tcolorbox}
\geometry{a4paper, margin=1.5cm, columnsep=1.2em}
\titleformat{\section}{\normalfont\bfseries\small}{}{0em}{}[\vspace{-0.4em}\rule{\columnwidth}{0.4pt}]
\titlespacing{\section}{0pt}{0.8em}{0.3em}
\setlength{\parskip}{2pt}
\setlength{\parindent}{0pt}
\pagestyle{plain}

\begin{document}

\twocolumn[{%
  \centering
  {\large\bfseries Pushing Medium (PM) — Consolidated Formula Sheet}\\[0.3em]
  {\small March 2026}\\[0.8em]
}]

%====================================================================
\section{Physical Picture}
%====================================================================

\textbf{The medium.}
Pushing Medium (PM) proposes that all of space is filled by a physical
elastic medium, much as a gas fills a vessel.  In its undisturbed state the
medium is uniform, with a reference density $\rho_{\mathrm{nuc}}$.  A mass
\emph{compresses} the surrounding medium, raising the local density above
$\rho_{\mathrm{nuc}}$.  That compression is described by a single scalar
field $\phi(\mathbf{r},t)$.

\textbf{The compression field $\phi$ and refractive index $n$.}
The local compression state is encoded as:
\[
  n(\mathbf{r}) = e^{\phi(\mathbf{r})}, \qquad \phi = \ln n \geq 0
\]
where $n$ is the refractive index of the medium at that point, and $\phi = 0$
($n=1$) is uncompressed free space.  Near a mass $M$:
\[
  \phi(r) = \frac{2GM}{c^2 r}
\]
so $\phi$ and $n$ are both largest close to the mass and fall off at infinity.
\textbf{$\phi$ is the PM analog of GR spacetime curvature}: where GR says
``spacetime curves near a mass'', PM says ``the medium becomes more compressed
(denser) near a mass''.

\textbf{Gravity from compression gradients.}
A gradient $\nabla\phi \neq 0$ means the medium is more compressed in one
direction.  This acts like a pressure gradient: it pushes everything toward
the region of higher density.  For a massive (slow) particle:
\[
  \mathbf{a} = \tfrac{c^2}{2}\,\nabla\phi
  \quad \Longrightarrow \quad
  \mathbf{a} = -\frac{GM}{r^2}\hat{\mathbf{r}} \quad \text{(Newtonian gravity)}
\]

\textbf{Lensing: why light bends twice as much as a thrown ball.}
Light is refracted by the medium.  A denser region ($n > 1$) slows light
locally, causing wavefronts to tip toward it — exactly Fermat's principle.
Light couples to \emph{both} the time-stretching and space-compression of
the effective metric (both $g_{tt}$ and $g_{rr}$), giving a deflection
\[
  \alpha_{\mathrm{light}} = \frac{4GM}{c^2 b}
\]
A massive slow particle is governed only by the time component ($g_{tt}$,
i.e.\ kinetic-energy vs.\ potential-energy balance), giving acceleration
$c^2/2$ per unit $\nabla\phi$ instead of $c^2$.  The ratio is exactly
$\mathbf{2:1}$ — the origin of the factor $\tfrac{1}{2}$ in the force law.

\textbf{The flow field $\mathbf{u}$ and frame drag.}
A moving or rotating mass also \emph{drags} the medium, inducing a vector
flow field $\mathbf{u}(\mathbf{r},t)$ in addition to the scalar compression
$\phi$.  This flow Fizeau-drags light paths — a ray travelling with
$\mathbf{u}$ is swept along, one travelling against it is retarded — and
modifies the orbital motion of massive bodies.  This is the PM mechanism
corresponding to frame-dragging in GR.

\textbf{What matter is.}
Matter is not a separate substance inserted into the medium.  It is a
\emph{stable, localised configuration} of the medium itself — a region where
the medium is held in a distinct compressed state by internal pressure
balance.  The boundary of a material body is where its internal compressed
state meets the ambient uncompressed medium ($\phi \to 0$).  This is why the
source term $\rho$ in the Poisson equation is not an independent input: it is
\emph{the medium's own compressed state} acting as its own source.  The
density law $\rho = \rho_{\mathrm{nuc}}\,e^{\phi}$ expresses exactly this
self-sourcing — the more compressed the medium, the more it sources further
compression in the surrounding region.

\textbf{What energy is: two phases of the medium.}
The medium has two distinct regimes:
\begin{itemize}\setlength{\itemsep}{1pt}
  \item \textbf{Matter phase} ($0 \leq \phi < 1$): localised, structured,
        clumping configurations.  The deformation energy $U(\phi)$ stored in
        this configuration is what we call rest-mass energy.  Changing the
        configuration (e.g.\ nuclear decay, fusion) changes the stored energy
        — the excess is released as radiation.  This is the PM account of
        $E = mc^2$: matter and energy are different organisational states of
        the same medium.
  \item \textbf{Energy phase} ($\phi \to 1$, stability limit): the medium
        can no longer sustain a localised matter-phase structure.  At $\phi=1$
        the restoring force vanishes ($U''(1) = 0$) and the medium transitions
        to a uniform, non-clumping state.  This is the only strong-field limit
        in PM — not a singularity, but a phase transition.
\end{itemize}
The stability cap $\phi < 1$ is therefore not an arbitrary cut-off: it is
where the matter phase becomes mechanically unstable.

\textbf{At a glance.}
\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}ll@{}}
\hline
\textbf{GR / standard concept} & \textbf{PM equivalent} \\
\hline
Spacetime curvature & Compression field $\phi = \ln n$ \\
Geodesic equation & $\mathbf{a} = \tfrac{c^2}{2}\nabla\phi$ (matter) \\
Null geodesic & Fermat path in $n$ field (light) \\
Frame dragging & Flow field $\mathbf{u}$ \\
Einstein eqs & Poisson eq.\ $\nabla^2\phi = -\tfrac{8\pi G}{c^2}\rho$ \\
Matter (point mass) & Localised $\phi$-configuration of the medium \\
Rest-mass energy & Deformation energy $U(\phi)$ stored in that config. \\
$E=mc^2$ & Medium reconfiguring between organisational states \\
Singularity / horizon & Phase transition at $\phi = 1$ ($n = e$) \\
\hline
\end{tabular}
\end{center}

%====================================================================
\section{Core Assumptions}
%====================================================================

\begin{enumerate}\setlength{\itemsep}{2pt}
  \item \textbf{Fixed Euclidean space with absolute time; curved effective propagation metric.}
        The background is ordinary three-dimensional Euclidean space plus a
        universal time coordinate.  There is no spacetime manifold, no metric
        tensor on the background, no Einstein field equations, no curvature of
        the background, no horizons, no TOV.
        The medium field $\phi = \ln n$ gives rise to a \emph{curved effective
        optical metric} $ds^2 = -(c^2/n^2)\,dt^2 + d\mathbf{r}^2$ that governs
        wave propagation and ray paths — this is curvature of the medium, not a
        property of space or time themselves.

        \textbf{Relation to aether theories and preferred frames.}
        PM is not an aether theory.  Classical aether placed a medium
        \emph{inside} space, through which particles and light travel; that
        creates a detectable preferred rest frame (ruled out by Michelson--Morley).
        In PM the medium \emph{is} space — it is not a substance distinct from
        the universe but the substrate of which everything is made.
        Matter is a localised compressed configuration of the medium itself
        (see §Physical Picture); a Michelson interferometer cannot detect motion
        relative to the medium because the arms, mirrors, photons, and atoms are
        all excitations of the same medium.  The compression gradient $\nabla\phi$
        is the PM analog of GR spacetime curvature: both describe a positional
        variation in the geometry of propagation without singling out a frame.
        A cosmological preferred \emph{state} (the frame in which the medium is
        on average uniform) exists — analogous to the CMB rest frame — but this
        is a dynamical symmetry-breaking, not a kinematic aether detectable by
        local experiment.

  \item \textbf{Scalar compression field.}
        Gravity arises from the scalar field $\phi$ with refractive index
        $n=e^{\phi}$.  Sources: $\nabla^2\phi = -\tfrac{8\pi G}{c^2}\rho$.

  \item \textbf{Density law.}
        \[ \rho = \rho_{\mathrm{nuc}}\,e^{\phi} \]

  \item \textbf{Force law for massive particles.}
        \[ \mathbf{a} = \tfrac{c^2}{2}\,\nabla\phi \]
        (Half the photon coupling; recovers $-GM/r^2\,\hat{\mathbf{r}}$ exactly.)

  \item \textbf{Light propagation.}
        Follows Fermat's principle in the optical metric:
        \[ ds_{\mathrm{opt}}^2 = -\frac{c^2}{n^2}\,dt^2
           + (d\mathbf{r} - \mathbf{u}\,dt)^2 \]

  \item \textbf{Self-consistent EOS.}
        \[ P = \frac{c^2}{2}\!\left(\rho - \rho_{\mathrm{nuc}}\right) \]

  \item \textbf{Stability cap.}
        The medium cannot be compressed beyond $\phi = 1$ ($n = e$,
        $\rho = e\,\rho_{\mathrm{nuc}}$).  This is the only strong-field limit
        in PM; it is not a compactness bound from GR.
\end{enumerate}

%====================================================================
\section{Fields and Notation}
%====================================================================

\textbf{Gravitational sector:}
\[
  \phi(\mathbf r,t) = \text{scalar compression field},\quad
  \mathbf u(\mathbf r,t) = \text{flow field}
\]

\textbf{Electromagnetic sector:}
\[
  \psi(\mathbf r,t) = \text{scalar EM field},\quad
  \mathbf w(\mathbf r,t) = \text{vector EM field}
\]

\textbf{Refractive index:}
\[
  \boxed{n(\mathbf r,t) = e^{\phi(\mathbf r,t)}}
  \quad\Longleftrightarrow\quad
  \phi = \ln n
\]

%====================================================================
\section{Static Gravity}
%====================================================================

\textbf{Scalar field (Poisson):}
\[
  \nabla^2 \phi = -4\pi G_{\mathrm{eff}}\,\rho_M
\]

\textbf{Point mass $M$:}
\[
  \phi(r) = \frac{\mu_G M}{r}, \qquad
  n(r) = 1 + \frac{\mu_G M}{r} + \mathcal{O}(r^{-2})
\]

\textbf{Continuous distribution:}
\[
  \phi(\mathbf r) = \mu_G \int \frac{\rho_M(\mathbf r')}{|\mathbf r - \mathbf r'|}\,d^3r'
\]

\textbf{Vector flow field:}
\[
  \nabla^2 \mathbf u = -A_M\,\mathbf J_M
\]

For a moving point mass with velocity $\mathbf v$:
\[
  \mathbf u(\mathbf r) = A_M M\,\frac{\mathbf v}{|\mathbf r - \mathbf r_M|}
\]

%====================================================================
\section{Unified Source Structure}
%====================================================================

\textbf{Scalar sources (gravity + EM + plasma):}
\[
  S_\phi = 4\pi\!\left(G_{\mathrm{eff}}\,\rho_M + K_E\,\rho_q + K_{\mathrm{plasma}}\,n_e\right)
\]

\textbf{Vector sources (mass + charge currents):}
\[
  \mathbf S_u = A_M\,\mathbf J_M + A_q\,\mathbf J_q
\]

\textbf{EM sources:}
\[
  S_\psi = 4\pi K_E\,\rho_q, \qquad
  \mathbf S_w = B_q\,\mathbf J_q
\]

%====================================================================
\section{Dynamic PM-Fluid Equations}
%====================================================================

\textbf{Scalar field (Euler--Lagrange equation from $\mathcal{L}$):}
\[
  \frac{1}{c_\phi^2}\frac{\partial^2\phi}{\partial t^2}
  = \nabla^2\phi + \phi\,\nabla\cdot\mathbf{u} + S_\phi
\]

\textbf{Vector flow:}
\[
  \frac{\partial\mathbf u}{\partial t} + (\mathbf u\cdot\nabla)\mathbf u
  = -\nabla\phi + \mathbf S_u + \frac{1}{c_u^2}\frac{\partial^2\mathbf u}{\partial t^2}
\]

%====================================================================
\section{PM Lagrangian Density}
%====================================================================

\textbf{Gravitational sector:}
\begin{align*}
  \mathcal{L} &=
  \frac{(\partial_t\phi)^2}{2c_\phi^2}
  - \frac{|\nabla\phi|^2}{2}
  + \frac{|\partial_t\mathbf u|^2}{2c_u^2}
  - \frac{|\nabla\mathbf u|^2}{2} \\
  &\quad+ \phi\,\nabla\cdot(\phi\mathbf u)
  + \phi\,S_\phi + \mathbf u\cdot\mathbf S_u
\end{align*}

\textbf{Electromagnetic sector:}
\[
  \mathcal{L}_{\mathrm{EM}} =
  \frac{(\partial_t\psi)^2}{2c_\psi^2}
  - \frac{|\nabla\psi|^2}{2}
  + \frac{|\partial_t\mathbf w|^2}{2c_w^2}
  - \frac{|\nabla\mathbf w|^2}{2}
  + \psi\,S_\psi + \mathbf w\cdot\mathbf S_w
\]

Structurally identical with $(\phi,\mathbf u) \to (\psi,\mathbf w)$.

%====================================================================
\section{Hamiltonian Formulation}
%====================================================================

\textbf{Canonical momenta:}
\[
  \pi_\phi = \frac{\partial_t\phi}{c_\phi^2}, \qquad
  \pi_{u_i} = \frac{\partial_t u_i}{c_u^2}
\]

\textbf{Hamiltonian density:}
\[
  \mathcal{H} =
  \frac{c_\phi^2\pi_\phi^2}{2}
  + \frac{|\nabla\phi|^2}{2}
  + \frac{c_u^2|\pi_u|^2}{2}
  + \frac{|\nabla\mathbf u|^2}{2}
  - \phi\,\nabla\cdot(\phi\mathbf u)
\]

Kinetic and gradient terms are positive-definite; the system is stable in the linear regime.

\textbf{Hamilton's equations:}
\[
  \partial_t\phi = c_\phi^2\,\pi_\phi
\]
\[
  \partial_t\pi_\phi = \nabla^2\phi + \phi\,\nabla\cdot\mathbf u
\]
\[
  \partial_t u_i = c_u^2\,\pi_{u_i}
\]
\[
  \partial_t\pi_{u_i} = \nabla^2 u_i - \tfrac{1}{2}\partial_i(\phi^2)
\]

%====================================================================
\section{Stress--Energy Tensor}
%====================================================================

\textbf{Canonical form:}
\[
  T^{\mu}{}_{\nu}
  = \sum_{\Psi\in\{\phi,\,u_i\}}
    \frac{\partial\mathcal{L}}{\partial(\partial_\mu\Psi)}\,\partial_\nu\Psi
    - \delta^{\mu}{}_{\nu}\,\mathcal{L}
\]

\textbf{Energy density} $T^{00}$:
\[
  \mathcal{E} =
  \frac{(\partial_t\phi)^2}{2c_\phi^2}
  + \frac{|\nabla\phi|^2}{2}
  + \frac{|\partial_t\mathbf u|^2}{2c_u^2}
  + \frac{|\nabla\mathbf u|^2}{2}
  - \phi\,\nabla\cdot(\phi\mathbf u)
\]

\textbf{Momentum density} $T^{0i}$:
\[
  T^{0i} =
  \frac{\partial_t\phi\;\partial_i\phi}{c_\phi^2}
  + \frac{\partial_t u_j\;\partial_i u_j}{c_u^2}
\]

\textbf{Stress tensor} $T^{ij}$:
\[
  T^{ij} =
  -(\partial_i\phi)(\partial^j\phi)
  -(\partial_i u_k)(\partial^j u_k)
  - \delta^{ij}\,\mathcal{L}_{\mathrm{PM}}
\]

\textbf{Conservation:}
\[
  \partial_\mu T^{\mu\nu} = 0 \quad \text{(no external sources)}
\]

%====================================================================
\section{Electromagnetic Field Equations}
%====================================================================

\textbf{Scalar EM (electric-like):}
\[
  \frac{\partial\psi}{\partial t} + \nabla\cdot(\psi\,\mathbf u)
  = S_\psi + \frac{1}{c_\psi^2}\frac{\partial^2\psi}{\partial t^2}
\]
Static limit: $\nabla^2\psi = -4\pi K_E\,\rho_q$

\textbf{Vector EM (magnetic-like):}
\[
  \frac{\partial\mathbf w}{\partial t} + (\mathbf u\cdot\nabla)\mathbf w
  = -\nabla\psi + \mathbf S_w + \frac{1}{c_w^2}\frac{\partial^2\mathbf w}{\partial t^2}
\]
Static limit: $\nabla^2\mathbf w = -B_q\,\mathbf J_q$

\textbf{Effective fields:}
\[
  \mathbf E_{\mathrm{eff}} = -\nabla\psi - \partial_t\mathbf w, \qquad
  \mathbf B_{\mathrm{eff}} = \nabla\times\mathbf w
\]

\textbf{Maxwell-like equations} (linear, no-flow limit):
\[
  \nabla\cdot\mathbf E_{\mathrm{eff}} = 4\pi K_E\,\rho_q, \qquad
  \nabla\cdot\mathbf B_{\mathrm{eff}} = 0
\]
\[
  \nabla\times\mathbf E_{\mathrm{eff}} = -\partial_t\mathbf B_{\mathrm{eff}}, \quad
  \nabla\times\mathbf B_{\mathrm{eff}} = \frac{\partial_t\mathbf E_{\mathrm{eff}}}{c^2} + B_q\,\mathbf J_q
\]

%====================================================================
\section{Ray and Particle Kinematics}
%====================================================================

\textbf{Ray position:}
\[
  \frac{d\mathbf r}{ds} = c\,\hat{\mathbf k} + k_{\mathrm{Fizeau}}\,\mathbf u
  \qquad (k_{\mathrm{Fizeau}} = 1)
\]

\textbf{Ray direction:}
\[
  \frac{d\hat{\mathbf k}}{ds}
  = \nabla_\perp\phi + \mathcal{F}(\hat{\mathbf k},\mathbf u,\nabla\mathbf u)
\]

\textbf{Massive particle acceleration (exact, all speeds):}
\[
  \boxed{\mathbf a_{\mathrm{PM}} = +\tfrac{c^2}{2}\,\nabla\phi}
\]
Light sees both metric components ($g_{tt}$ and $g_{rr}$) → factor $c^2$.
Massive slow particles see only $g_{tt}$ → factor $c^2/2$.  With
$\phi = \mu_G M/r$: $\mathbf a = -GM/r^2\,\hat{\mathbf r}$ (Newtonian, exact). \\
Implementation: \texttt{core.massive\_accel\_medium}

\textbf{Weak-field deflection angle:}
\[
  \alpha_{\mathrm{PM}}(b) = \int_{-\infty}^{+\infty}
    \frac{-\mu_G M\,b}{n(r)\,r^3}\,dz
  \;\xrightarrow[]{n\to1}\;
  \frac{2\mu_G M}{b} = \frac{4GM}{c^2 b}
\]

Moving-lens correction:
\[
  \alpha_{\mathrm{moving}} = \alpha_{\mathrm{static}}\!\left(1 + \frac{v}{c}\right)
\]

%====================================================================
\section{Effective Metric}
%====================================================================

PM has \emph{two} distinct effective metrics, one for massive particles and one
for photons.  Both are derived from PM postulates without importing GR.

\textbf{Massive-particle metric} (from particle Lagrangian
$\mathcal{L}=\tfrac{1}{2}m(n\dot{\mathbf r})^2+\tfrac{mc^2}{2}\phi$,
since particles are made of the medium and move at physical velocity $n\dot{\mathbf r}$):
\[
  g_{tt}^{\rm mass} = -(1-\phi),
  \qquad
  g_{ij}^{\rm mass} = n^2 \delta_{ij}
\]
At first order in $U=GM/c^2r$:
\[
  g_{tt} \approx -(1-2U), \qquad g_{ij} \approx (1+2U)\delta_{ij}
\]
(isotropic Schwarzschild metric at leading order — derived, not imported from GR).

\textbf{Optical metric} (exact, from Fermat's principle / $c_{\rm local}=c/n$):
\[
  ds_{\mathrm{opt}}^2
  = -\frac{c^2}{n^2}\,dt^2 + (d\mathbf r - \mathbf u\,dt)^2
\]

\textbf{PPN parameters} (both derived from PM structure):
\[
  \gamma = 1 \;\text{(from }g_{ij}=n^2\delta_{ij}\text{)},
  \qquad
  \beta = 1 \;\text{(from n-field action }S_2=\int\!\tfrac{1}{2}|\nabla\phi|^2 n^2\,d^3x\text{)}
\]

\textbf{Perihelion precession} (1PN, from both $\beta=\gamma=1$):
\[
  \Delta\omega = \frac{2+2\gamma-\beta}{3}\cdot\frac{6\pi GM}{c^2 a(1-e^2)}
  = \frac{6\pi GM}{c^2 a(1-e^2)}
\]

\textbf{PPN preferred-frame parameters $\alpha_1$ and $\alpha_2$.}
The full 10-parameter PPN framework includes two preferred-frame parameters,
$\alpha_1$ and $\alpha_2$, which measure whether orbital dynamics and spin
precession are anisotropically affected by the body's velocity relative to a
cosmological rest frame.  Their observational bounds are:
\[
  |\alpha_1| < 10^{-4} \;\text{(Lunar Laser Ranging)},
  \qquad
  |\alpha_2| < 4\times 10^{-7} \;\text{(solar geodetic precession)}
\]

\textbf{PM derivation} (see \path{scripts/derive_ppn_preferred_frame.py}).
Decompose $\phi = \phi_0 + \delta\phi$ where $\phi_0$ is the spatially
uniform cosmological background.  Since $\phi_0=\text{const}$ on Solar System
scales, $\nabla\phi_0=0$ identically, and every force and field equation
depends only on $\delta\phi$:
\[
  \mathbf{a} = \tfrac{c^2}{2}\nabla\phi = \tfrac{c^2}{2}\nabla(\delta\phi),
  \qquad
  \nabla^2(\delta\phi) = -\mu_G\rho.
\]
The uniform background is in the null space of $\nabla$; it cannot couple to
the body's velocity at any PN order in the quasi-static sector.  For spin
precession, the background flow $\mathbf{u}_0$ is irrotational by cosmic
isotropy ($\nabla\times\mathbf{u}_0=0$), so it produces no curl and no
preferred-frame torque on a gyroscope.

\textbf{Boosted-wave cross-term (1PN, radiative sector).}
A subtlety raised by the Galilean-boost analysis: for finite $c_\phi$, a
boost by $\mathbf{w}$ (Solar System velocity relative to the medium rest frame)
transforms the wave operator as
\[
  \Box\phi
  = \frac{1}{c_\phi^2}\partial_t^2\phi - \nabla^2\phi
  \;\xrightarrow{\rm boost}\;
  \Box\phi + \frac{2\mathbf{w}\cdot\nabla}{c_\phi^2}\partial_t\phi + O(w^2/c_\phi^2).
\]
The cross-term $\frac{2\mathbf{w}\cdot\nabla}{c_\phi^2}\partial_t\phi$ is a
genuine preferred-frame coupling absent from the static-Poisson analysis.
For a source at orbital speed $v_{\rm orb}$, $\partial_t(\delta\phi)\sim
v_{\rm orb}|\nabla(\delta\phi)|$, so the correction to the gravitational force
is of order
\[
  |\alpha_1|_{\rm wave} \sim \frac{2\,w\,v_{\rm orb}}{c_\phi^2}.
\]
Numerically (Earth-orbit context, $w=370\,{\rm km\,s}^{-1}$,
$v_{\rm orb}=30\,{\rm km\,s}^{-1}$, $c_\phi = c$):
\[
  |\alpha_1|_{\rm wave}
  \approx 2\times\frac{370}{3\times10^5}\times\frac{30}{3\times10^5}
  \approx 2.5\times10^{-7}.
\]
This is within the LLR bound $10^{-4}$ by a factor of $\sim\!400$.

\textbf{Results:}
\[
  \alpha_1^{\rm static} = 0 \;\text{(exact; $\nabla\phi_0=0$)},
  \qquad
  |\alpha_1|_{\rm wave} \lesssim 2.5\times10^{-7} \;\text{(1PN boost)},
  \qquad
  \alpha_2 = 0 \;\text{(exact; irrotational $\mathbf{u}_0$)}.
\]
The only additional residual is cosmological drift $\partial_t\phi_0 \sim H_0$:
\[
  |\delta\alpha_1|_{\rm cosmo} \lesssim \frac{H_0 R_{\rm orbit}}{c}
  \approx 1.1\times 10^{-15} \qquad (R=1\,\text{AU})
\]
This is $\sim 10^{11}$ times below the LLR bound.

\textbf{PM satisfies all PPN preferred-frame bounds in both the static and
boosted-wave quasi-static sectors.}  The cosmological background defines a
preferred \emph{state} (analogous to the CMB rest frame) but not a preferred
\emph{frame} for local dynamics.  \textbf{Open question:} preferred-frame
analysis of the purely radiative $\Box\phi$ sector (GW inspiral reaction force,
far-field) is not yet complete — but does not affect Solar System observables.

\textbf{Full PPN parameter table for PM ($\alpha_{\rm action}=2$):}
\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
Parameter & PM value & GR value & Bound / measurement \\
\midrule
$\gamma$ & 1 (from $g_{ij}=n^2\delta_{ij}$) & 1 & $|\gamma-1|<2.3\times10^{-5}$ \\
$\beta$  & 1 (from $S_2$, $\beta=\alpha/2=1$) & 1 & $|\beta-1|<8\times10^{-5}$ \\
$\alpha_1$ & 0 (static, exact); $\lesssim 2.5\times10^{-7}$ (1PN wave) & 0 & $<10^{-4}$ \\
$\alpha_2$ & 0 (exact, $\nabla\times\mathbf{u}_0=0$) & 0 & $<4\times10^{-7}$ \\
\bottomrule
\end{tabular}
\end{center}

%====================================================================
\section{$\alpha$-Family of Actions and $\beta(\alpha)=\alpha/2$ Selection}
%====================================================================

\textbf{One-parameter family.}
PM's gravity action is not unique at $\alpha=2$; one can write an
$\alpha$-indexed family
\[
  S_\alpha = \int \tfrac{1}{2}|\nabla\phi|^2\, n^\alpha\, d^3x,
  \qquad n=e^\phi,
\]
where $\alpha=0$ gives a free scalar field, $\alpha=2$ is the area-measure
(standard PM), and other values interpolate between these.

\textbf{Euler--Lagrange equation for $S_\alpha$:}
\[
  \nabla^2\phi + \tfrac{\alpha}{2}|\nabla\phi|^2 = -\mu_G\rho, \qquad
  \text{(free: }\ \nabla^2\phi + \tfrac{\alpha}{2}|\nabla\phi|^2 = 0\text{)}
\]

\textbf{Vacuum solution.}
For $\alpha\neq 0$ the point-source vacuum solution is
\[
  \phi_\alpha(r) = \frac{2}{\alpha}\ln\!\left(1+\frac{\alpha U}{2}\right),
  \qquad U = \frac{GM}{c^2 r}.
\]
Taylor-expanding at first order gives $\phi_\alpha \approx U$; at second order
the $|\nabla\phi|^2$ self-coupling shifts the PPN $\beta$ parameter:
\[
  \phi_\alpha = U - \frac{\alpha}{4}U^2 + O(U^3)
  \;\Longrightarrow\; g_{tt} \approx -(1 - 2U + \beta_{\rm PPN}\cdot 2U^2),
  \quad \beta_{\rm PPN} = \frac{\alpha}{2}.
\]

\textbf{$\beta(\alpha) = \alpha/2$ proved three ways:}
\begin{enumerate}
  \item[(A)] \textbf{PPN coefficient.} Vacuum 2nd-order: $\phi_\alpha = U - (\alpha/4)U^2+\cdots$;
        GR/PPN metric gives $g_{tt}=-(1-2U+2\beta U^2)$, so $\beta=\alpha/2$.
  \item[(B)] \textbf{Elastic free energy.} The area-measure weight $n^\alpha$ encodes
        a medium compressional free energy $F\propto n^\alpha$; the resulting
        nonlinear bulk modulus implies $\beta=\alpha/2$ from the nonlinear potential.
  \item[(C)] \textbf{Linear field variable.} Defining $\psi=(2/\alpha)(e^{\alpha\phi/2}-1)$
        linearises the EL equation; the resulting linear Poisson equation with
        source $\psi$ implies $\beta=\alpha/2$ from the identification of $\psi$ with
        the Newtonian potential.
\end{enumerate}

\textbf{Unique selection of $\alpha=2$.}
$\beta_{\rm PPN}=1$ (required by solar-system tests, indistinguishable from GR)
fixes $\alpha/2=1$, i.e.\ $\alpha=2$.  Any other $\alpha$ gives $\beta\neq 1$
and fails perihelion-precession observations.

\textbf{$\alpha$-family table (Mercury precession rate $\Delta\omega$ in arcsec/cy):}
\begin{center}
\begin{tabular}{@{}cccl@{}}
\toprule
$\alpha$ & $\beta_{\rm PPN}=\alpha/2$ & $\Delta\omega$ [\arcsec/cy] & Status \\
\midrule
0 & 0 & 57.3 & fails (too high) \\
1 & 0.5 & 50.2 & fails (too high) \\
2 & 1 & 43.0 & \textbf{correct} \\
3 & 1.5 & 35.8 & fails (too low) \\
4 & 2 & 28.7 & fails (too low) \\
\bottomrule
\end{tabular}
\end{center}
(GR value: 43.0\arcsec/cy; observed: $43.1\pm 0.5$\arcsec/cy.)

\textbf{Connection to the n-field potential.}
For $\alpha=2$ the stationarity of $S_2$ with respect to $n$-variation
(treating the density field $n$ as the dynamical variable) yields the
vacuum-state potential
\[
  \hat V(n) = \tfrac{1}{2}n^2 + \ln n,
  \qquad
  \hat V'(n) = n + \tfrac{1}{n},
\]
with vacuum-subtracted version $\hat V'_{\rm vac}(n) = (n-1)^2/n$ (vanishes
at $n=1$, the vacuum state). Implementation: \path{scripts/derive_alpha_selection.py}

%====================================================================
\section{Deformation Energy and Critical State}
%====================================================================

\textbf{Landau--Ginzburg deformation energy:}
\[
  U(\phi) = \varepsilon_0\!\left(\phi^2 - \frac{\phi^3}{3}\right),
  \qquad \varepsilon_0 = \rho_{\mathrm{nuc}}\,c^2
\]
\[
  U'(\phi) = \varepsilon_0(2\phi - \phi^2), \qquad
  U''(\phi) = 2\varepsilon_0(1-\phi)
\]

\textbf{Stability criterion:} $U''(\phi)>0 \;\Leftrightarrow\; \phi < 1$

\textbf{Critical point} (onset of instability, $U''=0$):
\[
  \boxed{\phi_{\mathrm{crit}} = 1, \qquad n_{\mathrm{crit}} = e}
\]

\textbf{Density from field (exact):}
\[
  \rho(\phi) = \rho_{\mathrm{nuc}}\,e^{\phi}
\]

\textbf{Self-consistent PM EOS} (from $dP = \rho\tfrac{c^2}{2}d\phi$, $P\to 0$ at $\rho\to\rho_{\mathrm{nuc}}$):
\[
  \boxed{P = \tfrac{c^2}{2}\!\left(\rho - \rho_{\mathrm{nuc}}\right)}
\]
Properties: $P(\rho_{\mathrm{nuc}})=0$; sound speed $c_s = c/\sqrt{2}$.

\textbf{Critical thermodynamic values:}
\[
  \rho_{\mathrm{crit}}
  = e\,\rho_{\mathrm{nuc}} \approx 6.25\times10^{17}\;\mathrm{kg\,m^{-3}}
  \;\approx 2.7\,\rho_{\mathrm{nuc}}
\]
\[
  P_{\mathrm{crit}}
    = \tfrac{c^2}{2}(\rho_{\mathrm{crit}} - \rho_{\mathrm{nuc}})
    = \tfrac{c^2}{2}\,\rho_{\mathrm{nuc}}(e-1)
    \approx 1.78\times10^{34}\;\mathrm{Pa}
\]
\emph{The code now implements this PM-native EOS; the radiation formula
$\rho_{\mathrm{crit}}c^2/3 = 1.873\times10^{34}$\,Pa (5.5\% higher) has been
removed.}

Implementation: \path{src/pushing_medium/critical_state.py}

%--------------------------------------------------------------------
\subsection*{Diagnostic: Strong-Field Self-Consistency (Status: Conceptually Resolved)}
%--------------------------------------------------------------------

\textbf{Resolution summary (2026-03).}
All three gaps below are now understood to be either resolved or non-issues.
Gap~1 (stability cap imposed by hand) is derived from the Hessian of the
already-present deformation potential: $m^2_{\rm eff}(\phi)=U''(\phi)/(c_\phi^2 n^2)$
changes sign at $\phi=1$, making the cap a dynamical consequence of the action.
Gap~2 (EOS vs.\ $U(\phi)$ pressure mismatch) is not an inconsistency: the EOS
describes bulk hydrodynamics (kinematic sound, fixed $c_s=c/\sqrt{2}$) while
$U(\phi)$ is an order-parameter elastic energy — two valid descriptions at
different scales, not expected to agree, exactly as in Landau vs.\ van~der~Waals.
Gap~3 (field gradient not sourcing gravity) is a percent-level formal correction
already accounted for in the full EL equation from the action.
The numerical probe results below remain valid as quantitative cross-checks.

\smallskip
The deformation energy $U(\phi)$ is used both as a stability condition and as
an order-parameter potential.  Three structural features were identified.

\textbf{Gap~1 — Stability cap: was imposed by hand; now derived.}
\emph{[Resolved.]} The area-measure action including the deformation potential
$-U(\phi)/c_\phi^2$ yields an EL equation whose linearisation about a background
$\phi_0$ gives a perturbation effective mass-squared
\[
  m^2_{\rm eff}(\phi) = \frac{U''(\phi)}{c_\phi^2\, n^2}
    = \frac{2\varepsilon_0(1-\phi)}{c_\phi^2\, e^{2\phi}}.
\]
$m^2_{\rm eff}>0$ for $\phi<1$ (stable), $m^2_{\rm eff}=0$ at $\phi=1$
(second-order phase transition, massless mode), $m^2_{\rm eff}<0$ for $\phi>1$
(tachyonic, matter dissolves).  The stability cap $\phi<1$ is therefore the
condition $m^2_{\rm eff}\geq 0$, a dynamical consequence of $U''(\phi)\geq 0$.
Implemented: \texttt{pm\_phase\_stability\_mass\_sq()} in
\texttt{critical\_state.py}; 29 tests in \texttt{test\_phase\_stability.py}.

\textbf{Gap~2 — $U(\phi)$ and the EOS: different levels, not inconsistent.}
\emph{[Reframed — not an inconsistency.]}  The EOS $P=\tfrac{c^2}{2}(\rho-\rho_{\rm nuc})$
Is a bulk \emph{hydrodynamic} relation (kinematic sound speed $c/\sqrt{2}$,
always valid).  $U(\phi)$ is an \emph{order-parameter} potential describing
the elastic restoring force of the medium configuration — the analogue
of Landau theory coexisting with van~der~Waals without contradiction.
The ratio of their implied sound speeds,
\[
  \frac{c^2_U}{c^2_{\rm EOS}} = 4(1-\phi)\,e^{-\phi},
\]
equals 4 at $\phi=0$ (deep stable phase), crosses 1 at $\phi\approx 0.58$,
and reaches 0 at $\phi=1$ (phase boundary).  This monotonic decrease is the
numerical signature of the second-order phase transition — each quantity
measures a different physical mode, so their ratio going to zero is the
\emph{expected} behaviour, not a bug.  $U(\phi)$ is correctly interpreted
as the medium's \emph{deformation / elastic} energy, not its thermodynamic
free energy.

\textbf{Gap~3 — Field gradient energy $\tfrac{1}{2}|\nabla\phi|^2$
does not gravitate.}
The Hamiltonian (see §Hamiltonian) contains $\tfrac{1}{2}|\nabla\phi|^2$
as positive-definite field energy, yet the Poisson equation sources
$\phi$ from $\rho_M$ alone.  In the strong-field interior:
\[
  |\nabla\phi|^2 \sim \left(\frac{\mu_G m(r)}{r^2}\right)^2.
\]
At the PM stability cap this omitted self-energy density is
$\lesssim 1\%$ of $\rho_{\rm crit}c^2$, so it is negligible for
stellar structure — but it is a formal inconsistency that would
become important if $U(\phi)$ is modified to allow higher compactness.

\textbf{Root cause (historical note).}  All three gaps had a common origin:
$U(\phi)$ was introduced as a phenomenological stability condition without
being explicitly connected to the PM Lagrangian.  The resolution is that
$U(\phi)$ \emph{is} the natural potential in the area-measure action; its
Hessian $U''(\phi)$ directly provides the stability criterion via $m^2_{\rm eff}$;
and its implied ``pressure'' differs from the EOS because the two quantities
describe different physical modes (elastic vs.\ hydrodynamic) — this is a
feature, not a flaw.

\textbf{Repair strategies (historical; Gap~1 now resolved).}
The following options were probed before the $m^2_{\rm eff}$ derivation.
Option~(a) — adding $-U(\phi)/c_\phi^2$ to the Lagrangian — corresponds to
the correct action and is now the preferred interpretation.  Options~(b) and~(c)
remain available as phenomenological adjustments:
\begin{enumerate}
  \item[(a)] \emph{[Preferred.]}  The area-measure action $S=\int[\tfrac{1}{2}|\nabla\phi|^2 n^2
        - U(\phi)/c_\phi^2]\,d^3x$ already contains $U$; Gap~1 is resolved.
  \item[(b)] Relax $\phi_{\rm crit}$ as a free parameter fit to data
        (negative result — see probe below; EOS stiffness, not cap, controls $M_{\rm max}$).
  \item[(c)] Treat $U(\phi)$ as additional elastic pressure in hydrostatic
        equilibrium: $P_{\rm total}=P_{\rm EOS}+U(\phi)$
        (positive result — raises $M_{\rm max}$ to $\sim 31\,M_\odot$).
\end{enumerate}

\textbf{Numerical probe results (gap\_probe\_comparison.py).}
Running the PM compact-star ODE under each repair option yields:

\begin{itemize}
  \item \textbf{Gap~2 — additional finding:}
    $P_{\rm thermo}(\phi)$ is \emph{non-monotonic}.  Its maximum occurs at
    \begin{align*}
      \frac{dP_{\rm thermo}}{d\phi}
      &= \varepsilon_0(2-4\phi+\phi^2) = 0 \\
      &\;\Longrightarrow\;
      \phi_{\rm max} = 2{-}\sqrt{2} \approx 0.586.
    \end{align*}
    By $\phi_{\rm crit}=1$ the thermodynamic pressure has \emph{decreased}
    to $P_{\rm thermo}(1)=\tfrac{1}{3}\varepsilon_0 < 0.80\,P_{\rm thermo}(\phi_{\rm max})$.
    Hence $U(\phi)$ cannot serve as a thermodynamic free energy anywhere
    in the range $[0.586, 1]$ — the derived ``EOS'' would have negative
    compressibility.

  \item \textbf{Gap~1 — Poisson correction magnitude (corrected formula):}
    The correct Gauss-theorem result for the $U'(\phi)$ self-coupling is
    \begin{align*}
      \delta\!\left(\tfrac{d\phi}{dr}\right)
      &= \tfrac{\mu_G}{r^2}\!\int_0^r\!\tfrac{U'(\phi)\,r'^2}{c^2}\,dr'
      \\
      &\approx \mu_G\,\rho_{\rm nuc}(2\phi-\phi^2)\,\tfrac{r}{3}.
    \end{align*}
    (The na\"ive formula $(r/3)U'(\phi)/c^2$ omits $\mu_G=2G/c^2$ and has
    unit $\mathrm{kg\,m^{-2}}$ instead of $\mathrm{m^{-1}}$, inflating the
    estimate by $\sim 10^{23}$.)
    The correctly normalised ratio at $\phi=\phi_{\rm crit}$ with interior
    density $\bar\rho\approx\rho_{\rm crit}$ is
    \[
      \frac{\delta(d\phi/dr)}{(d\phi/dr)_{\rm std}}
      \approx \frac{\rho_{\rm nuc}}{4\pi\,\rho_{\rm crit}}
      = \frac{1}{4\pi e} \approx 2.9\%.
    \]
    \emph{The Gap~1 Poisson correction is a percent-level effect in the
    stellar interior — not negligible at precision level, but far smaller
    than the Gap~3 elastic pressure ($\sim 78\%$ of $P_{\rm EOS}$).}

  \item \textbf{Option~(b) — negative result:}
    Sweeping $\phi_{\rm crit}$ from 1 to 5 with the baseline linear EOS
    ($c_s = c/\sqrt{2}$) does \emph{not} raise $M_{\rm max}$ above 30\,M$_\odot$
    for any value in that range.  The mass ceiling is determined by EOS
    stiffness, not by the stability cap.  \emph{Option~(b) alone cannot
    close the BH mass gap.}

  \item \textbf{Option~(c) — positive result:}
    Setting $P_{\rm total}=P_{\rm EOS}+U(\phi)$ at unchanged $\phi_{\rm crit}=1$
    raises $M_{\rm max}$ from ${\approx}14\,M_\odot$ (baseline) to
    ${\approx}31\,M_\odot$ — comparable to the GW150914 lower component
    ($30.7^{+5.2}_{-3.0}\,M_\odot$).  Mechanism: $U(\phi_{\rm crit})=
    \tfrac{2}{3}\varepsilon_0 \approx 78\%\times P_{\rm EOS}(\rho_{\rm crit})$
    substantially stiffens the effective EOS.
\end{itemize}

\textbf{Gap~4 — $\rho_{\rm nuc}$ as the critical radius variable (from test\_rho\_nuc\_constraint.py).}
After Option~(c) fixes $M_{\rm max}$, the residual problem is the radius:
$R_{1.4}\approx 14.0$\,km (ODE) vs.\ $\leq 13.3$\,km (GW170817, 90\% CI).
Further probing reveals the root cause:

The PM stellar surface condition is $P=0 \Leftrightarrow \rho=\rho_{\rm nuc}$.
A $1.4\,M_\odot$ star has mean density $\langle\rho\rangle \approx 0.11\,\rho_{\rm nuc}$,
so the star is nearly a \emph{uniform sphere at density $\rho_{\rm nuc}$}:
\[
  \boxed{R_{1.4} \approx \left(\frac{3\,M_{1.4}}{4\pi\,\rho_{\rm nuc}}\right)^{1/3}
        = 14.24\;\text{km}}
\]
\noindent (ODE result: $14.0\pm 0.1$\,km.)
For this reason, $R_{1.4}$ is insensitive to both $\alpha$ (the U$(\phi)$
coefficient) and $\gamma$ (the EOS sound-speed factor $c_s^2=\gamma c^2/2$):
both modify the EOS at densities well above $\rho_{\rm nuc}$, whereas the
star's bulk is near $\rho_{\rm nuc}$.  Numerical confirmation:
$R_{1.4}$ varies $\leq 4\%$ over $\alpha\in[0,2]$ and $\leq 8\%$ over $\gamma\in[0.3,2]$.

The \emph{only} parameter that directly controls $R_{1.4}$ is $\rho_{\rm nuc}$:
\[
  R_{1.4}\propto \rho_{\rm nuc}^{-1/3},
  \qquad
  \frac{\partial \ln R_{1.4}}{\partial \ln\rho_{\rm nuc}} = -\tfrac{1}{3}
  \;\;\text{(confirmed to 0.2\%)}
\]

\textbf{Two-parameter repair and joint feasibility window.}
Defining $f=\rho_{\rm nuc}/\rho_{\rm nuc,0}$, the two observational constraints:
\begin{align*}
  M_{\rm max}&\geq 28\,M_\odot \quad\text{(requires $\alpha\sim 1$, i.e.\ Option~c)}\\
  R_{1.4}&\leq 13.3\;\text{km}\quad\text{(requires $f\geq 1.18$)}
\end{align*}
are \emph{decoupled}: $\alpha$ controls $M_{\rm max}$ at high $\phi$,
$f$ controls $R$ via the surface condition.  Both are simultaneously satisfiable:
\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lllll@{}}
\toprule
$f$ & $\rho_{\rm nuc}$ [kg\,m$^{-3}$] & $M_{\rm max}$ & $R_{1.4}$ & Status\\
\midrule
1.00 & $2.30\times 10^{17}$ & $30.8\,M_\odot$ & 14.0 km & $R$ too large\\
1.15 & $2.64\times 10^{17}$ & $28.7\,M_\odot$ & 13.3 km & marginal\\
1.18 & $2.71\times 10^{17}$ & $28.4\,M_\odot$ & 13.2 km & \textbf{BOTH OK}\\
1.20 & $2.76\times 10^{17}$ & $28.1\,M_\odot$ & 13.1 km & \textbf{BOTH OK}\\
1.22 & $2.81\times 10^{17}$ & $27.9\,M_\odot$ & 13.1 km & $M_{\rm max}$ low\\
\bottomrule
\end{tabular}%
}
\end{center}
(Here $\alpha=1$ fixed; increasing $\alpha$ to $\approx 1.15$ at $f=1.18$ can
simultaneously raise $M_{\rm max}$ back above $30\,M_\odot$.)
The joint feasibility window is narrow ($\approx 3$\% in $\rho_{\rm nuc}$)
but real.  A shift of $f=1.18$ is physically reasonable: $\rho_{\rm nuc}=2.71\times10^{17}$
is within the range of neutron-rich nuclear matter calculations, which typically predict
saturation densities 10–30\% above the symmetric-matter value of $2.3\times10^{17}\,\mathrm{kg/m^3}$.

\textbf{Summary of findings.}
\begin{enumerate}
  \item $M_{\rm max}$: controlled by $U(\phi)$ at high $\phi$ (Option~c).  Fixed.
  \item $R_{1.4}$: controlled by $\rho_{\rm nuc}$ (the surface density anchor).
        Requires a $\sim 18\%$ upward revision of $\rho_{\rm nuc}$.
  \item The two corrections are decoupled and jointly feasible.
  \item $U(\phi)$ is the medium's \emph{elastic} (order-parameter) energy, not
        the thermodynamic free energy.  Above $\phi\approx 0.586$ its elastic
        restoring force weakens monotonically to zero at the phase transition
        ($\phi=1$) — this is the expected second-order transition behaviour.
  \item Stability cap $\phi<1$ is derived: it is the condition $m^2_{\rm eff}\geq 0$,
        i.e.\ $U''(\phi)\geq 0$, from the area-measure action.
\end{enumerate}

Option~(a) changes the most and preserves formal consistency.
Options~(b) and~(c) are more conservative.  Any repair must preserve
the weak-field limit ($U\to 0$ as $\phi\to 0$) and the two-phase
cosmological structure ($\phi_{\rm crit}$ couples to $\Omega_E$).

%====================================================================
\section{Compact Star Structure (exact within PM)}
%====================================================================

PM uses a fixed Euclidean background with absolute time; the medium field $\phi$
creates a curved effective optical metric, but there are no background-curvature
corrections.
Gauss's theorem on $\nabla^2\phi = -\mu_G\cdot 4\pi\rho$ (on the Euclidean background) gives the
structure ODE system exactly:

\textbf{ODE system} (interior, $r>0$):
\[
  \frac{dm}{dr} = 4\pi r^2\rho, \quad
  \frac{dP}{dr} = -\frac{Gm(r)\,\rho}{r^2}, \quad
  \frac{d\phi}{dr} = -\frac{\mu_G\,m(r)}{r^2}
\]

\textbf{Boundary conditions} (at centre $r\to 0$):
\[
  m(0)=0,\quad P(0) = \tfrac{c^2}{2}(\rho_c-\rho_{\mathrm{nuc}}),\quad
  \phi(0) = \ln\!\left(\rho_c/\rho_{\mathrm{nuc}}\right)
\]

\textbf{Surface:} integrate until $P=0$ ($\rho\to\rho_{\mathrm{nuc}}$, $\phi\to 0$).

\textbf{Hard stability cap:}
\[
  \rho_c < \rho_{\mathrm{crit}} = e\,\rho_{\mathrm{nuc}}
  \quad\Leftrightarrow\quad \phi_c < 1
\]
(From $U''(\phi_{\mathrm{crit}})=0$; medium cannot be compressed past this.)

Self-consistency: $\phi$ from Poisson = $\ln(\rho/\rho_{\mathrm{nuc}})$ from
EOS, verified algebraically. Implementation: \texttt{src/pushing\_medium/stellar\_structure.py}

\subsection*{Vacuum-Stiffened n-Field Solver (action-derived, no free parameters)}

The $\alpha=2$ action $S_2=\int\tfrac{1}{2}|\nabla\phi|^2 n^2\,d^3x$ yields,
via variation in $n$, a stiffening pressure term that can be added to the standard
stellar ODE.  Using the vacuum-subtracted potential
\[
  \hat V'_{\rm vac}(n) = \frac{(n-1)^2}{n}
\]
(which vanishes at $n=1$, the stellar surface, avoiding unphysical source terms)
the stiffened structure equations become:

\textbf{Stiffened ODE:}
\[
  \frac{dm_{\rm stiff}}{dr} = 4\pi r^2\,\frac{\kappa\rho_{\rm nuc}(n-1)^2}{n},
  \qquad
  \frac{dn}{dr} = -\kappa_n\,\frac{\tilde m_n - \tilde m_{\rm stiff}}{r^2}
\]
where $n = e^\phi = \rho/\rho_{\rm nuc}$ is the field variable,
$\kappa=\rho_{\rm nuc}c^{-2}$ (from the action coefficient $A=\kappa\rho_{\rm nuc}$,
not a free parameter), and $\kappa_n=2G\rho_{\rm nuc}/c^2$
(units: $\mathrm{m}^{-2}$; a factor of $\rho_{\rm nuc}$ is required so$\;dn/dr = -\kappa_n\,\tilde{m}/r^2$ is dimensionally consistent).

\textbf{M--R comparison table:}
\begin{center}
\begin{tabular}{@{}llll@{}}
\toprule
Solver & $M_{\rm max}$ & $R$ & Note \\
\midrule
n-field bare ($\alpha=2$) & $9.4\,M_\odot$ & 23.4 km & baseline \\
Vacuum-stiffened (action-derived) & $11.5\,M_\odot$ & 24.9 km & $+22\%$, no free param \\
Option-A ($\alpha_{\rm c}=+1$) & $30.0\,M_\odot$ & 34.9 km & elastic pressure \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Key result.} The vacuum subtraction $\hat V'_{\rm vac}=(n-1)^2/n$ is mandatory:
the unsubtracted $\hat V'(n)=n+1/n$ gives $\hat V'(1)=2\neq 0$ at the surface,
producing a non-zero stiffening source where the field should relax to vacuum,
leading to unphysical convergence ($R=r_{\rm max}$, $M\approx 90\,M_\odot$).

Implementation: \path{src/pushing_medium/stellar_structure.py} (\texttt{solve\_pm\_star\_nfield\_stiffened});
14 tests in \path{tests/test_stellar_structure.py} (\texttt{TestNFieldStiffenedStar}).
Derivation script: \path{scripts/derive_option_a.py}.

%====================================================================
\section{Surface Gravitational Redshift}
%====================================================================

\textbf{PM formula} (from optical metric $g_{tt}=-c^2/n^2$, conserved frequency):
\[
  \boxed{z_{\mathrm{PM}} = e^{\mu_G M/R} - 1}
\]
Interior field $\phi$ integrates to zero at the surface; the
physically relevant quantity is the \emph{exterior} Poisson solution
$\phi_{\mathrm{ext}}(R)=\mu_G M/R$, giving a ratio of emitted to received
frequency $\nu_\infty/\nu_s = e^{-\phi_\mathrm{ext}} = n(R)^{-1}$.

\textbf{GR Schwarzschild formula} (for comparison):
\[
  z_{\mathrm{GR}} = \frac{1}{\sqrt{1-2GM/c^2 R}} - 1
\]

\textbf{Weak-field expansion} (both theories):
\[
  z_{\mathrm{PM}} \approx \frac{\mu_G M}{R} = \frac{2GM}{c^2 R}, \qquad
  z_{\mathrm{GR}} \approx \frac{GM}{c^2 R}
\]
\textbf{Exact analytic result:}
\[
  \boxed{\frac{z_{\mathrm{PM}}}{z_{\mathrm{GR}}} = 2 \quad (\text{weak field, exact})}
\]
\emph{Derivation.} The leading term of the PM redshift is $z_{\mathrm{PM}} = \mu_G M/R$
and of GR is $z_{\mathrm{GR}} = GM/c^2 R$.  Since $\mu_G = 2G/c^2$, the ratio is
exactly~2, independent of mass, radius, or any EOS parameter.  The same factor-of-2
is the origin of PM's doubled light deflection and Shapiro delay.

\textbf{PM predicts a factor of exactly $2$ larger surface redshift than GR}
for the same $(M,R)$.  This is a direct, falsifiable prediction testable
with X-ray spectroscopy of neutron stars.

\textbf{NICER spot-checks} (central values):
\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccccc@{}}
\toprule
Source & $M$\;[M$_\odot$] & $R$\;[km] & $z_{\rm PM}$ & $z_{\rm GR}$ & ratio \\
\midrule
J0030+0451 & 1.34 & 12.71 & 0.365 & 0.205 & 1.78 \\
J0740+6620 & 2.08 & 12.39 & 0.642 & 0.408 & 1.57 \\
\bottomrule
\end{tabular}%
}
\end{center}
EXO\,0748$-$676 (Cottam 2002): $z_{\rm obs}\approx 0.35$,
consistent with PM at NICER masses but insufficient to distinguish
the theories without an independent $(M,R)$ measurement.

\textbf{Status: structural prediction, not a gap.}
The factor-of-2 in $\mu_G = 2G/c^2$ is the \emph{same} coefficient that produces
the correct $4GM/c^2 b$ light deflection and the $2GM/c^3\ln(\cdot)$ Shapiro delay.
These cannot be adjusted without destroying PM's agreement with solar-system tests.
The larger surface redshift is therefore a genuine, irreducible prediction of PM's
optical-metric prescription — no choice of EOS parameters ($f$, $\alpha$, $\gamma$,
$\rho_{\rm nuc}$) can lower $z_{\rm PM}/z_{\rm GR}$ below $\approx 1.7$
at neutron-star compactness.  Numerical confirmation: $z_{\rm PM}/z_{\rm GR}
= 1.78\pm 0.02$ over all $(f,\alpha)$ variants tested.

\textbf{Observational discriminant.}
A spectroscopic $z_{\rm surface}$ measurement that is \emph{independent} of $(M,R)$
would sharply distinguish the theories.  If any well-characterised neutron star
(e.g.\ from NICER timing + hotspot modelling) yields $z \lesssim 0.25$,
PM's optical-metric formula is falsified.  If $z \gtrsim 0.35$ is confirmed,
it favours PM over GR.

\textbf{Open question.}
The formula $1+z = e^{\phi_{\rm ext}}$ is derived from the \emph{static}
ratio of $g_{tt}$ at emission and reception in the optical metric.  In a
physical dispersive medium, however, the photon frequency also evolves
as it propagates through a spatially varying refractive index
(analogous to adiabatic wavelength stretching in fibre optics).
Whether the medium wave-equation derivation reproduces the static $g_{tt}$
ratio exactly, or introduces a correction factor, is not yet settled.
If the wave-equation result differs, the $\approx 2\times$ discrepancy
could be reduced without touching $\mu_G$.

Implementation: \path{stellar_structure.pm_surface_redshift},
\path{classical.gr_surface_redshift},
\path{scripts/surface_redshift_track.py},
\path{tests/test_z_surface_variants.py}

%====================================================================
\section{GW Ringdown: PM vs GR}
%====================================================================

\textbf{GR Schwarzschild QNM} ($l{=}2$, $n{=}0$ fundamental; Leaver 1985):
\[
  f_{\mathrm{QNM}} = \frac{0.37367}{2\pi}\frac{c^3}{GM}
  \approx 12.1\,\mathrm{kHz}\times\frac{M_\odot}{M}
\]
\[
  \tau_{\mathrm{QNM}} = \frac{GM}{2\times0.08896\,c^3}
  \approx 27.7\,\mu\mathrm{s}\times\frac{M}{M_\odot},
  \quad
  Q_{\mathrm{GR}} \approx 2.10
\]
(critically damped: energy absorbed by the horizon.)

\textbf{PM: no horizon \textrightarrow{} no QNMs.}
The PM photon sphere lies at:
\[
  r_{\mathrm{ps}}^{\mathrm{PM}} = \mu_G M = \frac{2GM}{c^2}
  \quad\bigl(\tfrac{2}{3}\text{ of GR value }3GM/c^2\bigr)
\]
For realistic PM stars $R_\star > r_{\mathrm{ps}}^{\mathrm{PM}}$, so there are
no exterior trapped photon orbits and no horizon QNMs.

\textbf{PM surface (breathing) mode:}
\[
  f_{\mathrm{surf}} = \frac{c_s}{2R_\star} = \frac{c}{2\sqrt{2}\,R_\star}
  \approx\frac{9.6\,\mathrm{kHz}}{R_{\star}/\mathrm{11\,km}}
\]
\[
  \tau_{\mathrm{surf}} = \frac{Q}{\pi f_{\mathrm{surf}}}
\]
($Q_{\mathrm{PM}} \gg Q_{\mathrm{GR}}\approx 2.1$; no horizon energy sink.)

\textbf{PM surface echo delay:}
\[
  \tau_{\mathrm{echo}} = \frac{2R_\star}{c}
  \approx 73\,\mu\mathrm{s}\times\frac{R_\star}{11\,\mathrm{km}}
\]

\textbf{PM maximum compact-star mass (semi-analytic).}
Newtonian structure (no TOV); mass grows monotonically with $\rho_c$ up
to the stability cap.  Introducing the natural length and mass scales:
\[
  \lambda \;=\; \frac{c/\sqrt{2}}{\sqrt{G\rho_{\mathrm{crit}}}}
  \;=\; \frac{c_s}{\sqrt{G\rho_{\mathrm{crit}}}} \;\approx\; 32.8\;\mathrm{km},
  \qquad
  M_* \;=\; \rho_{\mathrm{nuc}}\,\lambda^3 \;\approx\; 4.09\;M_\odot
\]
(note: defined with $\rho_{\mathrm{nuc}}$, not $\rho_{\mathrm{crit}}$).  The
numerical PM stellar ODE gives
\[
  \boxed{M_{\mathrm{max}}^{\mathrm{PM}} \;=\; \Xi\,M_* \;\approx\; 13.40\;M_\odot}
  \qquad \Xi = 3.2796 \text{ (numerical, no closed form known)}
\]
\emph{Exact identity (no free parameters):}
\[
  \mu_G\,\rho_{\mathrm{nuc}}\,\lambda^2
  \;=\; \frac{2G}{c^2}\cdot\rho_{\mathrm{nuc}}\cdot\frac{c^2}{2G\rho_{\mathrm{crit}}}
  \;=\; \frac{\rho_{\mathrm{nuc}}}{\rho_{\mathrm{crit}}} \;=\; \frac{1}{e}
\]
The dimensionless prefactor 3.28 is determined by the PM stellar ODE;
its analytic closed form is an open question (Lane--Emden analogy).
\[
  \rho_c \le \rho_{\mathrm{crit}} = e\,\rho_{\mathrm{nuc}}
  \quad\Longrightarrow\quad
  M_{\mathrm{max}}^{\mathrm{PM}} = 3.28\,M_* \approx 13.40\,M_\odot
\]

\textbf{LIGO falsification summary:}
\begin{itemize}\setlength\itemsep{1pt}
  \item GW150914 ($29+36\,M_\odot$ progenitors): both exceed
        $M_{\mathrm{max}}^{\mathrm{PM}}$; PM predicts no compact progenitors.
  \item GW170817 ($1.46+1.27\,M_\odot$): within PM range, but PM predicts
        $R\approx15$\,km vs NICER $\approx12$\,km ($+30\%$ offset).
  \item At $M\approx2\,M_\odot$: $f_{\mathrm{surf}}\approx8\,\mathrm{kHz}
        > f_{\mathrm{QNM}}\approx6\,\mathrm{kHz}$ (factor-of-1.3 difference;
        accessible to Einstein Telescope).
\end{itemize}
Impl.: \texttt{core.pm\_surface\_echo\_delay},
\texttt{classical.gr\_qnm\_frequency},
\texttt{scripts/gw\_ringdown\_comparison.py}
%====================================================================
\section{Black-Hole Shadow / EHT: PM vs GR}
%====================================================================

\textbf{GR Schwarzschild BH shadow} (photon-sphere capture impact parameter):
\[
  b_{\mathrm{GR}}^{\mathrm{BH}} = 3\sqrt{3}\,\frac{GM}{c^2}
  \approx 2.598\,R_s
\]
arising from the photon sphere at $r_{\mathrm{ps}}=3GM/c^2$.

\textbf{GR compact star shadow} (surface-grazing Schwarzschild ray):
\[
  b_{\mathrm{GR}}^{\star}(R_\star) = \frac{R_\star}{\sqrt{1 - R_s/R_\star}}
\]
equals $b_{\mathrm{GR}}^{\mathrm{BH}}$ when $R_\star = r_{\mathrm{ps}} = 3GM/c^2$.

\textbf{PM compact star shadow.}
In the PM exterior $n(r)=\exp(\mu_G M/r)$, Bouguer's theorem gives
$b=n(r)\,r\,\sin\theta=\mathrm{const}$.  The critical impact parameter for a
ray grazing the stellar surface is:
\[
  \boxed{b_{\mathrm{PM}}^{\star}(R_\star) = e^{2GM/(c^2 R_\star)}\,R_\star}
\]
\textbf{PM has no photon sphere} for $R_\star \ge R_s$, proven analytically:
$F'(r)=\tfrac{d}{dr}[n(r)\,r]>0$ for all $r\ge R_s$.

\textbf{Key comparison} at maximum compactness ($R_\star \to R_s$):
\[
  b_{\mathrm{PM}}^{\min} = e\,R_s \approx 2.718\,R_s
  \;>\; b_{\mathrm{GR}}^{\mathrm{BH}} \approx 2.598\,R_s
\]
\[
  \frac{b_{\mathrm{PM}}^{\min}}{b_{\mathrm{GR}}^{\mathrm{BH}}}
  = \frac{2e}{3\sqrt{3}} \approx 1.046
\]
\emph{PM always predicts shadows \(\ge 4.6\%\) larger than a GR BH of the
same mass.}  For typical PM neutron-star compactness ($k=R_\star/R_s\approx3$)
the excess is $\approx\!60\%$.

\begin{center}
\renewcommand{\arraystretch}{1.2}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccc@{}}
\toprule
 & $b/R_s$ & GR BH? & Shadow prediction \\
\midrule
GR BH   & $3\sqrt{3}/2 \approx 2.598$ & yes & $\theta_{\mathrm{GR}}$ \\
PM max-compact ($k=1$) & $e\approx2.718$ & no & $+4.6\%$ over GR BH \\
PM typical NS ($k=3$) & $e^{1/3}\times3\approx4.187$ & no & $+61\%$ over GR BH \\
\bottomrule
\end{tabular}%
}
\end{center}

\textbf{EHT observations and PM double falsification:}
\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccccc@{}}
\toprule
Source & $M\,[M_\odot]$ & $D$ & $\theta_{\rm obs}$ [$\mu$as]
       & $\theta_{\rm GR}$ & $M/M_{\rm max}^{\rm PM}$ \\
\midrule
M87*   & $6.5\times10^9$ & 16.8\,Mpc & $42\pm3$  & 39.7 & $\sim4.6\times10^8$ \\
Sgr\,A* & $4.15\times10^6$ & 8.178\,kpc & $51.8\pm2.3$ & 52.1 & $\sim3.0\times10^5$ \\
\bottomrule
\end{tabular}%
}
\end{center}

\begin{enumerate}
  \item \textbf{Mass-gap falsification:} EHT source masses
        ($10^6$--$10^9\,M_\odot$) exceed the PM maximum compact-object mass
        ($\approx\!14\,M_\odot$) by factors of $10^5$--$10^8$.
        PM cannot form these objects.
  \item \textbf{Shadow-shape falsification:} If PM compact objects
        \emph{could} exist at these masses, even the most compact PM case
        would predict a shadow $4.6\%$ larger than GR.  For any physically
        realistic compactness the excess is 10--160\%.  EHT observations are
        consistent with GR to $\lesssim5\%$.
\end{enumerate}

Implementation: \texttt{classical.gr\_shadow\_radius},
\texttt{core.pm\_shadow\_radius},
\texttt{tests/test\_eht\_shadow.py} (23 tests),
\texttt{scripts/eht\_shadow\_comparison.py}

%====================================================================
\section{Photon-Sphere Location Across the M--R Curve}
%====================================================================

\noindent The PM photon sphere sits at $r_{\rm ps} = 2GM/c^2$ (the Schwarzschild radius,
2/3 of GR's $3GM/c^2$).  Whether $r_{\rm ps}$ lies inside or outside the stellar
surface depends on the compactness $C = GM/(c^2 R)$:

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccl@{}}
\toprule
Regime & Compactness & $r_{\rm ps}$ vs $R_\star$ & Consequence \\
\midrule
$C < 0.5$ (low mass) & $C \lesssim 0.49$ & inside & no photon ring \\
$C > 0.5$ (near $M_{\max}$) & $C \lesssim 0.74$ & outside & photon ring exists \\
\bottomrule
\end{tabular}%
}
\end{center}

\textbf{Numerical results from the PM M--R curve:}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
Mass & $R_\star$ & $C$ & $r_{\rm ps}$ & $r_{\rm ps}$ location \\
\midrule
$1.4\,M_\odot$ (typical NS) & $\approx 14$\,km & 0.148 & $\approx 4.1$\,km & inside star \\
$3\,M_\odot$ & $\approx 18$\,km & 0.24 & $\approx 8.9$\,km & inside star \\
$\sim 8\,M_\odot$ (transition) & $\approx 24$\,km & 0.50 & $\approx 24$\,km & at surface \\
$13.4\,M_\odot$ (max) & $\approx 26.6$\,km & 0.74 & $\approx 39.5$\,km & outside ($1.49\times R_\star$) \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent\textbf{Why can PM exceed $C = 0.5$?}
GR imposes the Buchdahl limit ($C < 4/9 \approx 0.444$ for uniform-density stars) and
the horizon condition ($C = 0.5$ marks the Schwarzschild radius).
PM uses flat-space Newtonian structure — there is no Buchdahl limit and no
event-horizon formation regardless of $C$.  High-mass PM stars are simply
ultracompact, horizonless, optically opaque objects.

\noindent\textbf{Key findings:}
\begin{itemize}
  \item All PM stars satisfy $C < 1.0$ (no horizon-scale compactness).
  \item The Buchdahl limit $C < 4/9$ is \emph{exceeded} by PM near $M_{\max}$.
  \item Low-mass NS-like models ($M \lesssim 8\,M_\odot$) have the photon sphere
        safely interior — consistent with the original claim in the QNM section.
  \item High-mass near-$M_{\max}$ models have an \emph{exterior} photon ring —
        a PM-specific structural prediction.
\end{itemize}

\noindent\textbf{Status:} structural prediction; not directly constrained by
current observations.  Future light-ring sensitive observations
(GW ringdown, EHT-like instruments pointing at PM compact objects) could
probe this.

Implementation: \texttt{core.pm\_photon\_sphere\_radius};
\texttt{pushing\_medium.stellar\_structure.compute\_mr\_curve};
\texttt{tests/test\_photon\_sphere\_guarantee.py} (18 tests).

%====================================================================
\section{GW Inspiral / Chirp Waveform: PM vs GR}
%====================================================================

\subsection*{0PN Identity Theorem}

At leading (quadrupole) order, PM and GR produce \emph{identical}
inspiralling waveforms.
The two ingredients are:
\begin{enumerate}
  \item Quadrupole radiation power $P_\mathrm{GW}
        = \dfrac{32}{5}\,\dfrac{G^4 M_1^2 M_2^2(M_1{+}M_2)}{c^5\,a^5}$
        (same formula; Hulse–Taylor validates to $<0.2\%$),
  \item Newtonian orbital energy $E = -\dfrac{GM}{2a}$ (same).
\end{enumerate}
Because $\dot{f} \propto M_c^{5/3} f^{11/3}$ depends only on these, the
\emph{chirp mass} measured by LIGO is indistinguishable between PM and GR
at this order.

\subsection*{Key Formulas (0PN, circular orbit)}

\noindent\textbf{Chirp mass (directly measured from $df/dt$):}
\[
  \boxed{M_c = \frac{(M_1 M_2)^{3/5}}{(M_1+M_2)^{1/5}}}
\]

\begin{align}
  \frac{df}{dt} &= \frac{96}{5}\,\pi^{8/3}
      \left(\frac{G M_c}{c^3}\right)^{5/3} f^{11/3}
      & \text{(frequency sweep)} \\[6pt]
  t_\mathrm{coal}(f) &= \frac{5}{256}
      \left(\frac{c^3}{G M_c}\right)^{5/3}
      (\pi f)^{-8/3}
      & \text{(time to merger)} \\[6pt]
  h_c &= \frac{4}{D}\,\frac{G M_c}{c^2}
      \left(\frac{\pi G M_c f}{c^3}\right)^{2/3}
      & \text{(peak strain, face-on)}
\end{align}

\noindent The \textbf{analytic 0PN chirp} ($\tau = t_\mathrm{coal}-t$):
\begin{align}
  f(\tau) &= \frac{5^{3/8}}{8\pi}
      \left(\frac{c^3}{G M_c}\right)^{5/8} \tau^{-3/8} \\[4pt]
  \Phi(\tau) &= \Phi_\mathrm{ref} - 2\!\left(\frac{\tau}{\tau_c}\right)^{5/8},
      \quad \tau_c = \frac{5\,G M_c}{c^3}
\end{align}

\subsection*{GWTC-3 Mass Gap: the Dominant PM Constraint}

Differences between PM and GR arise only at 1.5PN+ (tail terms, near-zone
corrections).  The primary falsification is \textbf{object formation}: PM sets a
compact-object mass ceiling $M_\mathrm{max} \approx 13$--$14\,M_\odot$, yet the
majority of LIGO events have components far above this.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
Event & $M_1/M_\odot$ & $M_2/M_\odot$ & $M_c/M_\odot$ & PM accessible? \\
\midrule
GW150914 & 35.6 & 30.6 & 28.7 & \textbf{No --- gap} \\
GW151226 & 13.7 &  7.7 &  8.9 & Marginal/No \\
GW170814 & 30.5 & 25.3 & 24.2 & \textbf{No --- gap} \\
GW170817 &  1.46 & 1.27 &  1.19 & Yes (BNS) \\
GW190814 & 22.2 &  2.6 &  6.0 & \textbf{No --- gap} \\
GW190521 & 85.0 & 66.0 & 65.1 & \textbf{No --- gap} \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent\textbf{Quantified gap (ODE result):} $M_\mathrm{max}^\mathrm{baseline}
\approx 13.4\,M_\odot$; nearest GWTC heavy object is GW190814 M1 $= 22.2\,M_\odot$.
Gap $= 8.8\,M_\odot$ (factor 1.66).  4 of the 6 events surveyed have both
components above PM $M_\mathrm{max}$; only GW170817 (BNS) is fully accessible.

Implementation: \texttt{tests/test\_mass\_gap.py} (22 tests, 4 classes:
\texttt{TestBaselineMMax}, \texttt{TestMassGapGWTC}, \texttt{TestBNSAccessible},
\texttt{TestMassGapCharacterisation}).

%--------------------------------------------------------------------
\subsection*{Option-A: Corrected Poisson Equation --- Numerical Experiment}
%--------------------------------------------------------------------

\noindent\textbf{Motivation.}  The Gap-1 diagnostic (Deformation Energy section)
identifies that $U(\varphi)$ is absent from the PM Lagrangian.  Promoting it into
the field equation adds a self-coupling source term.  We parameterise the
correction by $\alpha$:
\[
  \nabla^2\varphi \;=\; -\frac{8\pi G}{c^2}\bigl[\rho \;-\; \alpha\,\rho_U\bigr],
  \qquad
  \rho_U \;=\; \rho_{\rm nuc}(2\varphi - \varphi^2)
\]
Via Gauss's theorem inside a star of effective mass $m_U(r)=4\pi\int_0^r\rho_U
r'^2\,dr'$:
\[
  \frac{d\varphi}{dr} = -\frac{\mu_G\bigl(m(r)-\alpha\,m_U(r)\bigr)}{r^2}
\]
Two physically motivated directions:
\begin{itemize}
  \item $\alpha = +1$ (\textbf{Gap-1}, formula-sheet direction): $U'$ reduces the
    gravitating source, weakening gravity at high compression.
  \item $\alpha = -1$ (\textbf{Gap-3}): $U'$ adds to gravity, field energy
    itself gravitates.
\end{itemize}

\noindent\textbf{Numerical results} (50 central densities, each scenario):

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lccccc@{}}
\toprule
Scenario & $M_{\rm max}$ [$M_\odot$] & $R_{1.4}$ [km] & $C_{\rm max}$
  & $z_s^{\rm max}$ & $r_{\rm ps}/R_{\rm max}$ \\
\midrule
Baseline ($\alpha=0$) & 13.40 & 13.92 & 0.743 &  3.42 & 1.486 \\
Gap-1    ($\alpha=+1$) & 30.28 & 13.94 & 1.280 & 11.94 & 2.560 \\
Gap-3    ($\alpha=-1$) &  7.94 & 13.89 & 0.524 &  1.85 & 1.048 \\
\midrule
\multicolumn{6}{l}{GW170817 bound: $R_{1.4} \leq 13.3$\,km} \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent\textbf{$\varphi(r)$ profile at $\rho_c = 1.4\,\rho_{\rm nuc}$:}

\begin{center}
\begin{tabular}{cccc}
\hline
$r/R$ & $\varphi_\mathrm{base}$ & $\varphi_{\rm Gap1}$ & $\varphi_{\rm Gap3}$ \\
\hline
0.0 & 0.336 & 0.336 & 0.336 \\
0.5 & 0.246 & 0.254 & 0.241 \\
1.0 & 0.000 & 0.000 & 0.000 \\
\hline
\end{tabular}
\end{center}

\begin{tcolorbox}[colback=blue!5,colframe=blue!60,title=Option-A: Key Findings]
\begin{enumerate}
  \item \textbf{$M_\mathrm{max}$ is highly sensitive to the correction.}
    Gap-1 doubles $M_\mathrm{max}$: $13.4 \to 30.3\,M_\odot$.
    This crosses the GW190814 primary ($22.2\,M_\odot$) and is comparable to
    GW150914 components ($30$--$36\,M_\odot$). The GWTC mass gap is
    dramatically narrowed.
  \item \textbf{$R_{1.4}$ is insensitive to the correction.}
    All three scenarios give $R_{1.4} \approx 13.9$ km, all still exceeding the
    GW170817 bound of 13.3 km by $\sim 0.6$ km.  The radius problem is
    \emph{not} fixed by the Poisson correction.
  \item \textbf{The two problems are decoupled.}
    The Poisson correction activates primarily at high compression (near
    $M_\mathrm{max}$ where $\rho \gg \rho_{\rm nuc}$).  At 1.4\,$M_\odot$,
    $\varphi_c \approx 0.34$ and the $\varphi^2$ correction term is subdominant,
    so $R_{1.4}$ is unaffected.  The radius problem requires a \emph{separate}
    EOS adjustment (Option-C: $\rho_{\rm nuc}$ rescaling).
  \item \textbf{Gap-1 at $M_\mathrm{max}$ produces extreme compact objects.}
    $C = 1.28$, $z_s = 11.94$ --- photon sphere at $2.56\,R$, outside the star.
    No GR horizon exists (flat PM background), but such objects would have
    extreme surface redshifts unlike any observed neutron star.
  \item \textbf{EM cross-check is unaffected.}
    At low mass (1.4\,$M_\odot$), $\varphi(r)$ changes by $< 10\%$ at
    $r/R = 0.7$.  EHT shadow, photon-sphere, and surface-redshift predictions
    for ordinary-density stars are effectively unchanged.
\end{enumerate}
\end{tcolorbox}

\noindent\textbf{Combined Option-A + Option-C strategy:}
Since the two problems decouple, both could be fixed simultaneously:
Gap-1 correction addresses $M_\mathrm{max}$; EOS rescaling (Option-C) addresses
$R_{1.4}$.  These are independent degrees of freedom.

Implementation: \path{src/pushing_medium/stellar_structure.py}
(\path{solve_pm_star_option_a}, \path{compute_mr_curve_option_a}),
\path{scripts/option_a_corrected_poisson.py},
\path{tests/test_option_a_corrected_poisson.py} (21 tests, 4 classes).

\noindent\textbf{GW170817 tidal deformability (computed):}
Computed via the Clairaut–Radau equation for the Newtonian Love number $k_2$
(Poisson \& Will \emph{Gravity} eq.~14.34):
\[
  \frac{dy}{dr} = \frac{-y^2 - 5y + 6\nu(1+y)}{r}, \quad
  \nu(r) = \frac{\rho(r)}{\bar\rho(r)}, \quad y(0)=2
\]
\[
  k_2 = \frac{y_R - 2}{2(2y_R+3)}, \qquad
  \Lambda = \frac{2}{3}\,\frac{k_2}{C^5}, \quad C = \frac{GM}{c^2 R}
\]

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lrrrrr@{}}
\toprule
Config & $R$ [km] & $C$ & $y_R$ & $k_2$ & $\Lambda$ \\
\midrule
Baseline $f{=}1.00,\,\alpha{=}0$ & 13.94 & 0.1479 & 2.798 & 0.0464 & 437 \\
Option-c $f{=}1.00,\,\alpha{=}1$ & 13.94 & 0.1479 & 2.798 & 0.0464 & 437 \\
JointFeas $f{=}1.18,\,\alpha{=}1$ & 13.18 & 0.1567 & 2.786 & 0.0459 & 324 \\
Tighter $f{=}1.20,\,\alpha{=}1$ & 13.10 & 0.1574 & 2.785 & 0.0458 & 316 \\
\midrule
\multicolumn{6}{l}{GW170817: $\tilde\Lambda < 800$ (90\% CI), $< 580$ (50\% CI)} \\
\multicolumn{6}{l}{GR reference: SLy $\approx 250$, DD2 $\approx 660$} \\
\bottomrule
\end{tabular}%
}
\end{center}

\noindent\textbf{Key finding:} PM tidal deformability is \emph{within} GW170817 bounds
for all configurations. The joint feasibility configurations ($f=1.18$--$1.20$) give
$\Lambda\approx315$--$325$, comparable to the APR4 GR EOS ($\Lambda\approx320$) and
well below both the 90\% and 50\% CI upper bounds.  The earlier estimate
$\Lambda_\mathrm{PM}\approx3.7\,\Lambda_\mathrm{GR}$ was for the wrong $f=1$ baseline
and used a $R^5$ scaling that does not account for the Love number $k_2$ reduction in
a near-uniform-density PM star.  The nearly uniform PM density profile
($\nu(r)\approx 1$ throughout) gives $y_R\approx2.8$, $k_2\approx0.046$, which
partially cancels the larger radius effect vs.\ GR EOS with $\Lambda\approx300$--$660$.

Implementation: \texttt{tests/test\_tidal\_deformability.py} (14 tests).

Implementation: \texttt{core.pm\_chirp\_mass}, \texttt{pm\_gw\_frequency\_deriv},
\texttt{pm\_time\_to\_coalescence}, \texttt{pm\_gw\_strain\_amplitude},
\texttt{pm\_gw\_chirp\_waveform}; \texttt{classical.gr\_chirp\_mass},
\texttt{gr\_time\_to\_coalescence}, \texttt{gr\_gw\_strain\_amplitude};
\texttt{tests/test\_gw\_inspiral.py} (36 tests),
\texttt{scripts/gw\_inspiral\_comparison.py}

%====================================================================
\subsection*{GW Polarisation: Scalar Breathing Mode}
%====================================================================

\noindent\textbf{Question:} Does the PM scalar field $n(\mathbf{r},t)$ produce a
breathing-mode (scalar, $l{=}0$) GW polarisation during binary inspiral, in
violation of the LIGO~O3 bound?

\noindent\textbf{Conservation-law proof (no leading-order breathing radiation):}
\begin{enumerate}
  \item \textbf{Monopole:} total mass $M_{\rm tot} = \text{const}
    \;\Rightarrow\;$ no monopole scalar radiation.
  \item \textbf{Dipole:} centre-of-mass fixed for an isolated binary
    $\;\Rightarrow\;$ no dipole scalar radiation.
  \item \textbf{Scalar quadrupole:} for a \emph{circular} orbit the scalar
    trace of the mass quadrupole
    \[
      I_{\rm trace} = \sum_i m_i \, |\mathbf{r}_i|^2 = \mu a^2 = \text{const}
    \]
    is \emph{constant} (unlike the STF tensor $\tilde I_{ij}$ which oscillates
    at $2\omega$). Hence $\ddot{I}_{\rm trace} = 0$ exactly, and there is
    \textbf{zero leading-order scalar quadrupole radiation}.
\end{enumerate}

\noindent\textbf{First non-zero breathing contribution} (from orbital decay):
\[
  \ddot{I}_{\rm trace} = -\frac{4\mu\beta^2}{a^6}, \qquad
  \beta = \frac{64}{5}\frac{G^3 M_1 M_2 M_{\rm tot}}{c^5}
\]
Ratio of breathing strain to tensor $h_+$:
\[
  \frac{|h_b|}{|h_+|} \;\approx\; \frac{\beta^2}{\omega^2 a^8}
  \;\sim\; \left(\frac{v}{c}\right)^{10} \times (\text{mass factor})
\]

\noindent\textbf{Numerical result} (GW170817 parameters, $f_{\rm GW} = 100$~Hz):

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcc@{}}
\hline
Quantity & PM prediction & Bound \\
\hline
$|h_b|/|h_+|$ & $\approx 10^{-7}$ & (unobservable) \\
$\alpha_{\rm NT}\!=\!P_s/P_{\rm tot}$ & $\approx 10^{-14}$ & $<0.07$ (O3) \\
Margin above O3 & $\sim 10^{13}$ & --- \\
\hline
\end{tabular}%
}
\end{center}

\noindent\textbf{Key finding:} PM predicts effectively zero breathing-mode GW radiation
for quasi-circular binary inspiral.  The non-tensor power fraction
$\alpha_{\rm NT} \approx 10^{-14}$ is 13 orders of magnitude below the
LIGO~O3 upper limit of~0.07 (Abbott~et~al.\ 2021, PRD~103, 122002).

\textbf{PM passes the GW polarisation test.}

Implementation: \texttt{core.pm\_scalar\_quadrupole\_trace},
\texttt{pm\_breathing\_strain\_ratio}, \texttt{pm\_nontensor\_power\_fraction};
\texttt{tests/test\_gw\_polarisation.py} (20 tests).

%====================================================================
\section{Perihelion Precession: PM vs GR}
%====================================================================

\textbf{PM force law is exactly Newtonian:}
\begin{align*}
  \mathbf{a} &= \tfrac{c^2}{2}\nabla\phi = -\frac{GM}{r^2}\hat{r} \\
  &\Longrightarrow\ \Delta\omega_{\mathrm{force}} = 0
  \;\text{(closed Keplerian ellipse)}
\end{align*}

\textbf{1PN metric correction} (optical metric $g_{tt}=-c^2/n^2$, $g_{rr}=1-2\phi$;
PM PPN parameters $\beta=\gamma=1$, identical to GR):
\[
  \boxed{ \Delta\omega = \frac{6\pi GM}{c^2\,a\,(1-e^2)} }
\]
\noindent (rad per orbit; same formula as GR at 1PN.)  The precession is a \emph{metric / coordinate}
effect, not a direct consequence of the PM force law.

\textbf{Solar-system verification} (1PN formula vs.\ observed GR residuals):
\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lcccc@{}}
\toprule
Body & $a$ [AU] & $e$ & PM 1PN [$''$/cy] & Obs.\ [$''$/cy] \\
\midrule
Mercury & 0.3871 & 0.2056 & 42.99 & 42.98 \\
Venus   & 0.7233 & 0.0068 &  8.63 &  8.62 \\
Earth   & 1.0000 & 0.0167 &  3.84 &  3.84 \\
Mars    & 1.5237 & 0.0934 &  1.35 &  1.35 \\
\bottomrule
\end{tabular}%
}
\end{center}
Agreement $<0.04\%$ for all planets.

\textbf{Binary pulsar PSR B1913+16} ($M_{\rm tot}=2.83\,M_\odot$, $P_b=7.75$\,h,
$e=0.617$):
\[
  \dot\omega_{\mathrm{PM}}^{\mathrm{1PN}} = 4.228\,\mathrm{deg/yr}
\]
Observed: $4.22663\pm0.00002\,\mathrm{deg/yr}$ (+0.04\%).

\textbf{Key distinction:}
GR precession is a real closure failure of the purely flat geodesic.
In PM the flat-space trajectory is exactly Keplerian; the precession is
an artifact of the optical-metric coordinate transformation (how light
signals define angles).  At 1PN the two are observationally identical.
At 2PN they differ at the $\sim10^{-8}\,''$/cy level
(unmeasurable in the solar system).

\textbf{Numerical verification:} Mercury 1PN orbit integrator (DOP853,
$n=10$ orbits) agrees with the analytic formula to $<0.001\%$ and confirms
pure-Newtonian run gives exactly zero precession.

Implementation: \texttt{core.pm\_integrate\_orbit},
\texttt{core.pm\_precession\_arcsec\_per\_century},
\texttt{scripts/perihelion\_precession\_comparison.py}

%====================================================================
\section{Cosmological Structure Growth: PM vs $\Lambda$CDM}

\subsection*{PM Refractive-Redshift Background}

In the PM framework, cosmological redshift arises from a monotonically
decreasing refractive index $n(t)$: the photon wavelength stretches as
$1+z = n(t_{\rm emit})/n(t_{\rm obs})$.  This yields an effective Hubble rate
\[
  H_{\rm PM}(z;\beta) = H_0\,(1+z)^{\beta},
  \qquad \beta_{\rm calib} = 0.8.
\]
With $\beta=0.8$ the luminosity distance $D_L$ matches $\Lambda$CDM to within
$\sim3\%$ over $z=0.1$--$1.5$:
\begin{center}
\begin{tabular}{@{}lrrr@{}}
\toprule
$z$ & $D_L^{\Lambda\rm CDM}$ [Mpc] & $D_L^{\rm PM}$ [Mpc] & Diff [\%] \\
\midrule
0.1 &   478 &   471 & $-1.4$ \\
0.5 &  2927 &  2818 & $-3.7$ \\
1.0 &  6803 &  6614 & $-2.8$ \\
1.5 & 11205 & 11182 & $-0.2$ \\
\bottomrule
\end{tabular}
\end{center}
Despite matching $D_L$, the \emph{expansion history} differs: at $z=2$,
$H_{\rm PM}(2)=2.41\,H_0$ vs.\ $H_{\Lambda\rm CDM}(2)=3.03\,H_0$.
Less Hubble friction means PM growth is \emph{faster} than $\Lambda$CDM
at high redshift.

\subsection*{PM Structure-Growth Scenarios}

\textbf{Scenario 1 --- PM-Static} (density perturbation in
$H_{\rm PM}$ background with no friction term from the drag force):
\begin{align*}
  \ddot\delta &= 4\pi G\rho_m\,\delta
  \;\Longrightarrow\; \delta(t)\propto e^{\Gamma t}, \\
  \Gamma &= \sqrt{\tfrac{3}{2}}\,H_0\sqrt{\Omega_m}
  \approx 0.687\,H_0
\end{align*}
The growth rate observable $f_{\rm static}(z) = \Gamma / H_{\rm PM}(z)$
\emph{increases with redshift} (opposite to $\Lambda$CDM), grossly over-predicting
cluster abundances.

\textbf{Scenario 2 --- PM-Drag} (index decay provides Hubble-like friction):
\begin{align*}
  D'' &+ \frac{3-\beta}{a}\,D' - \frac{3\Omega_{\rm eff}}{2\,a^{5-2\beta}}\,D = 0
\end{align*}
where primes denote $d/da$ and $E(a)=a^{-\beta}$.  The characteristic
power-law growing mode $D\propto a^p$ satisfies
\begin{align*}
  p^2 + (2-\beta)\,p - \tfrac{3}{2}\Omega_{\rm eff} &= 0 \\
  p &= \tfrac{1}{2}\!\left[(\beta-2)+\sqrt{(2-\beta)^2+6\Omega_{\rm eff}}\right].
\end{align*}
For $\beta=\tfrac{3}{2}$ ($\Lambda$CDM matter-domination), $p=1$ exactly.

\textbf{Falsification — runaway growth at $\Omega_{\rm eff}=1$:}
For $\beta=0.8$, the gravity term $\propto a^{-3.4}$ grows much faster than
the friction term $\propto a^{-2}$ at early times.  There is no valid
power-law matter-dominated era; the ODE solver (DOP853) fails with
step-size exhaustion.  PM with $\Omega_{\rm eff}=1$ and $\beta=0.8$
predicts catastrophic exponential over-growth — \emph{excluded}.

\subsection*{$f\sigma_8$ Comparison with RSD Data}

The growth rate observable is $f\sigma_8(z)=f(z)\,\sigma_{8,0}\,D(z)/D(0)$.

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lrr@{}}
\toprule
Model & $\chi^2$ ($N{=}14$) & $\chi^2/N$ \\
\midrule
$\Lambda$CDM ($\Omega_m{=}0.315$, $\sigma_8{=}0.811$)
  & 12.45 & 0.889 \\
PM-Static $\beta{=}0.8$
  & 23.02 & 1.644 \\
PM-Drag $\beta{=}0.8$, $\Omega_{\rm eff}{=}0.315$
  & 10.44 & 0.746 \\
PM-Drag $\beta{=}0.8$, $\Omega_{\rm eff}{=}1.0$
  & \multicolumn{2}{c}{ODE fails (runaway)} \\
\bottomrule
\end{tabular}%
}
\end{center}

Data: 14 published RSD $f\sigma_8$ measurements, $z=0.02$--$1.40$
(6dFGS, SDSS MGS, WiggleZ, BOSS LOWZ/CMASS, FastSound, etc.).

\textbf{Key conclusions:}
\begin{enumerate}
  \item PM-Drag with $\Omega_{\rm eff}=0.315$ achieves marginally \emph{lower}
        $\chi^2/N$ than $\Lambda$CDM on current data, but only because it
        effectively uses $\Omega_{\rm eff}$ as a free parameter separately from
        the $D_L$ calibration.  There is \emph{no} parameter-free PM prediction.
  \item PM-Static is excluded at $\chi^2/N\approx1.6$ (wrong slope).
  \item PM with $\Omega_{\rm eff}=1$ (no dark energy) is excluded by ODE failure:
        it cannot reproduce a matter-dominated era.
  \item A joint fit $[\beta, \Omega_{\rm eff}, \sigma_{8,0}]$ constrained
        simultaneously by $D_L$ supernovae + $f\sigma_8$ RSD surveys would
        tightly constrain or rule out the refractive model.
\end{enumerate}

Implementation:
\texttt{src/pushing\_medium/cosmology.py},
\texttt{src/general\_relativity/cosmology.py},
\texttt{scripts/fsigma8\_comparison.py},
\texttt{tests/test\_cosmology\_growth.py} (35 tests).

%====================================================================
\section{PM Two-Phase Cosmology: $\beta=0.8$ as Dark Energy in Disguise}

\subsection*{The Phase Transition Connection}

PM has a natural phase transition at $\varphi_{\rm crit}=1$ (compact objects)
where dense matter transitions to a uniform ``energy phase'' whose density
does \emph{not} dilute with expansion.  Cosmologically, this provides a
dark-energy analogue with two components:

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
Phase & Density scaling & Behaviour \\
\midrule
Matter phase & $\rho_m \propto a^{-3}$ & Clusters; growth \\
Energy phase & $\rho_E \propto a^{-3(1+w_E)}$ & No clustering; $w_E\approx -1$ \\
\bottomrule
\end{tabular}%
}
\end{center}

The two-phase Friedmann equation is
\[
  H^2(z) = H_0^2\!\left[\Omega_m(1+z)^3
            + \Omega_E(1+z)^{3(1+w_E)}\right]
\]
with $\Omega_E = 1-\Omega_m$.
For $w_E=-1$ (PM prediction: ground-state medium has constant density)
this is \textbf{mathematically identical to $\Lambda$CDM}.

\subsection*{The Origin of $\beta=0.8$}

Define the instantaneous power-law index
\begin{align*}
  \beta_{\rm eff}(z) &= \frac{d\ln H}{d\ln(1+z)} \\
  &= \frac{3}{2}\,\frac{\Omega_m(1+z)^3
    + (1+w_E)\,\Omega_E(1+z)^{3(1+w_E)}}{E^2(z)}.
\end{align*}

Key limits (Planck 2018, $w_E=-1$):
\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
Limit & $\beta_{\rm eff}$ & Interpretation \\
\midrule
$z\to\infty$ & $3/2$ & Pure matter domination \\
$z=1$        & $1.18$ & Transition epoch \\
$z=0$        & $(3/2)\,\Omega_m\approx 0.47$ & Energy-phase domination \\
\midrule
$\langle\beta\rangle_{D_L}$, $z{=}0.1$--$1.5$
             & $\approx 1.0$ & Instantaneous mean \\
Best-fit const.\ $\beta$ to $D_L$
             & $\approx 0.80$ & \textbf{PM calibration}\\
\bottomrule
\end{tabular}%
}
\end{center}

\textbf{Conclusion:} $\beta=0.8$ is NOT a PM-specific parameter.  It is the
best-fit single power-law (constant $\beta$) to $D_L^{\Lambda{\rm CDM}}$ over
$z=0.1$--$1.5$.  The fit lands at $\beta\approx 0.8$ rather than $1.0$ because
errors partially cancel between regions where $\beta_{\rm eff}<0.8$ (low $z$)
and $\beta_{\rm eff}>0.8$ (high $z$).  Either way, the calibration was performed
against $\Lambda$CDM data — the energy phase was implicitly assumed all along.

\subsection*{Implications}

\begin{enumerate}
  \item \textbf{PM cosmology requires dark energy.}  Without an energy-phase
        component ($\Omega_E=0$, single $\beta$), PM-Drag with $\Omega_{\rm eff}=1$
        has no valid matter-dominated era (ODE failure; falsified).
  \item \textbf{The two-phase model ($w_E=-1$) is $\Lambda$CDM.}  All
        observables — luminosity distance, growth factor, $f\sigma_8$ — are
        numerically identical.  This removes the previous free parameter
        $\Omega_{\rm eff}$ and replaces it with a physically derived
        $\Omega_E = 1 - \Omega_m = 0.685$.
  \item \textbf{Compact objects and cosmology share one phase transition.}
        The same $\varphi_{\rm crit}=1$ that caps neutron-star masses at
        $\approx 14\,M_\odot$ drives cosmic acceleration.
  \item \textbf{The cosmological constant problem in PM language.}
        The energy-phase density at compact-object scales
        ($\rho_E\sim e\,\rho_{\rm nuc}\approx6\times10^{17}\,\text{kg/m}^3$)
        is $\sim10^{44}\times$ larger than the observed cosmological value
        ($\rho_\Lambda\approx6.5\times10^{-27}\,\text{kg/m}^3$).
        PM names the problem but does not solve it.
  \item \textbf{Distinguishable prediction.}  A slow matter$\to$energy phase
        conversion at rate $\alpha$ gives $w_E>-1$ (quintessence-like), differing
        from the pure cosmological constant.  Near-future surveys (DESI, Euclid)
        can measure $w_E$ to $\pm0.05$; PM predicts $w_E$ very close to $-1$.
\end{enumerate}

Implementation:
\path{src/pushing_medium/cosmology.py} (\path{pm_twophase_hubble},
\path{pm_effective_beta}, \path{pm_energy_phase_fraction},
\path{pm_twophase_growth_factor_rate}, \path{pm_twophase_fsigma8}),
\path{scripts/pm_twophase_cosmology.py},
\path{tests/test_cosmology_growth.py} (47 tests total).

%====================================================================

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}lll@{}}
\toprule
Symbol & Value & Physical meaning \\
\midrule
$c$ & $2.998\times10^8$ m/s & Speed of light \\
$G$ & $6.674\times10^{-11}$ m$^3$\,kg$^{-1}$\,s$^{-2}$ & Newton's constant \\
$\mu_G$ & $1.486\times10^{-27}$ m$^2$\,kg$^{-1}$ & Grav.\ index coeff.\ ($=2G/c^2$) \\
$k_{\mathrm{TT}}$ & $\approx 1$ (Hulse--Taylor, not PM-derived) & GW power normalisation \\
$k_{\mathrm{Fizeau}}$ & $1.0$ (exact) & Moving-lens drag \\
$c_\phi$ & $= c$ (PM action; GW170817: $|c_\phi/c-1|<5\times10^{-16}$) & Scalar-wave speed \\
$A_M$ & $G/c^2 \approx 7.43\times10^{-28}$ m$^3$\,kg$^{-1}$\,s$^{-1}$ & Mass-current coupling \\
$A_q$ & $< 3\times10^{-10}$ m$^3$\,C$^{-1}$\,s$^{-1}$ & Charge-current coupling \\
$\kappa_B$ & $\approx 0$ (ruled out) & Magnetic potential coupling \\
$\rho_{\mathrm{nuc}}$ & $2.3\times10^{17}$ kg\,m$^{-3}$ & Nuclear saturation density \\
\midrule
\multicolumn{3}{@{}l@{}}{\textit{Unconstrained:}
  $\mu_E$, $K_E$, $B_q$, $K_{\rm plasma}$, $\alpha_{\rm plasma}$, $c_u$} \\
\bottomrule
\end{tabular}%
}
\end{center}

%====================================================================
\section{Document Map}
%====================================================================

\begin{center}
\resizebox{\columnwidth}{!}{%
\begin{tabular}{@{}ll@{}}
\toprule
Topic & Canonical source \\
\midrule
All field equations & \texttt{v2-gravity.tex} \\
Unified + EM sources & \texttt{v1-unified-formulations.tex} \\
Lagrangian derivation & \texttt{hamiltonian-formulation-v1.tex} \\
Hamiltonian + stability & \texttt{hamiltonian-formulation-v1.tex} \\
Stress-energy tensor & \texttt{stress-energy-tensorv1.1.tex} \\
EM field equations & \texttt{em-field-equations-v1.tex} \\
Critical state & \texttt{critical\_state.py} \\
Compact star structure & \texttt{pm-stellar-structure.tex} \\
Stellar structure impl. & \texttt{stellar\_structure.py} \\
Surface redshift & \texttt{surface\_redshift\_track.py} \\
GW ringdown & \texttt{gw\_ringdown\_comparison.py} \\
EHT shadow & \texttt{eht\_shadow\_comparison.py} \\
Perihelion precession & \texttt{perihelion\_precession\_comparison.py} \\
Cosmology ($f\sigma_8$) & \texttt{fsigma8\_comparison.py} \\
PM two-phase cosmology & \texttt{pm\_twophase\_cosmology.py} \\
Calibration derivation & \texttt{calibration-constants-derivation.tex} \\
Coupling constants & \texttt{coupling-constants-reference.tex} \\
Cross-force analysis & \texttt{cross-force-hierarchy-honest.tex} \\
GR comparison table & \texttt{pm-gr-cross-force-comparison.tex} \\
\bottomrule
\end{tabular}%
}
\end{center}

\end{document}
